GPU-Beschleunigung der Material Point Method

### 2018-11-22 mscthesis.tex
discretization space

= \vec{\nabla}} \cdot \boldsymbol{P}(_0\boldsymbol{x},t) + {_0\boldsymbol{f}} ^ {\text{body}}(_0\boldsymbol{x},t) {^t_0J}.

@@

\subsubsection{Weak formulation}

@@

Any conservative integrator could be used to discretize time. The implicit midpoint rule is one of such ($\lambda =\frac{1}{2}$)

@@

  \frac{y ^ { n + 1 } - y ^ { n }}{ \Delta t} =  f \left(t^n+\frac{\Delta t}{2}, \frac { y ^ { n } + y^{n+1}}{2}\right)

@@

@@

>>>>

@@

\subsubsection{Weak formulation}\label{sec:weak}

@@

Any conservative integrator could be used to discretize time. The implicit midpoint rule ($\lambda =\frac{1}{2}$) is one of such

@@

  \frac{y ^ { n + 1 } - y ^ { n }}{ \Delta t} = f^{n+\lambda} = f \left(t^n+\lambda\Delta t, (1-\lambda) y ^ { n } +\lambda y^{n+1}\right)

@@

\begin{equation}\label{eq:time_disc}

@@

\subsubsection{Discretize space}
A Galerkin discretization brings all spatial terms of equation \ref{eq:time_disc} to a finite-dimensional space: $\boldsymbol{q} \rightarrow \boldsymbol{q}_h$. To not further clutter up the notation, the $h$ will be omitted. This will replace $q_\alpha, v^n_\alpha , v_\alpha^{n+1}$ with their finite-dimensional grid-based interpolants:
  {_t q } _ { \alpha } ^ {  { n } } =  ({_t q } _ {  { i } \alpha } ^ {  { n } })  ({_t w } _ { i }) , \quad _tv _ { \alpha } ^ {  { n } } = ({_tv _ {  { j\alpha } } ^ {  { n } }} )( {_t w } _ {  { j } }) , \quad {_tv _ { \alpha } ^ {  { n } + 1 }} = ({_tv _ {  { j } \alpha } ^ {  { n } + 1 }}) ({_t w } _ {  { j } }).
Further chapter \ref{sec:weak} mentions that $\boldsymbol{q}$ can be chosen arbitrarily. The Galerkin discretization of a $d$-dimensional space with $m$ grid nodes therefore uses the standard basis functions $e_1,e_2, ... ,e_{d \times m}$. Due to the scalar-product $d\times m$ equations would need to be solved:
\begin{equation}\label{eq:galerkin}
  \frac{1}{\Delta t} \int _{ \Omega^{t^n}} ({ _tw } _ {i  })(_t\rho)({_t w } _ { j  })({_tv_{j\alpha}^{n+1}}-{_tv_{j\alpha}^{n}}) d_t\boldsymbol{x} =
 - \int _ { \Omega ^ {  t^ n } }   \frac{\partial{_t w } _ { {i  }}}{\partial_tx_\beta }  \sigma _ {{ \alpha } \beta }  { d } \boldsymbol { x }.
A mass matrix can be factored out as:
  m_{ij}^n = \int _{ \Omega^{t^n}} {_t w } _ {i  }(_t\rho){_t w } _ { j  } d_t\boldsymbol{x}.
The Lagrangian pull-back relates this to the initial density in the Lagrangian view and discretizing the integral with the initial time-invariant particle mass $m_p \approx V_p^0\rho(_0x_p,0)$:
m_{ij}^n = \int _{ \Omega^{t^0}} ({_tw_i})(_0\rho_0)({_tw} _ { j  }) d_0 \boldsymbol{x} \approx \sum _ {  { p } }  { m } _ {  { p } }  { w } _ {  { i } } (\boldsymbol { x } _ {  { p } } )  { w } _ {  { j } } ( \boldsymbol { x } _ {  { p } } ).
This matrix is symmetric positive semi-definite (since mass is positive). Numerically this matrix is mostly not used as is due to possibility of it being singular. This is solved commonly due to a mass lumping strategy. Replace $m^n_{ii}$ with the $i$-th row sum to get:
\begin{equation}\label{eq:mass_discr}
\sum _ {  { p } }  { m } _ {  { p } }  { w } _ {ip}   w_{jp}
\stackrel{\text{\ref{eq:partition_unity}}}{\approx}
\sum_p m_p w_{ip}
m_i,
where partition of unity $\sum_j w_{jp} = 1$ is used. This is exactly the mass transfer as in eq. \ref{eq:mpm:mass}, s.t. no further assembling of a mass matrix is needed.
Discretizing the right side of \ref{eq:galerkin} with an estimated per particle stress $\boldsymbol{\sigma}_p^{n+\lambda}$:
\begin{equation}\label{eq:stress_discr}
\int _ { \Omega ^ {  t^ n } }   \frac{\partial{_t w } _ { {i  }}}{\partial_tx_\beta }  \sigma _ {{ \alpha } \beta }  { d } \boldsymbol { x } \approx
\sum _ { p } (\sigma _ { p }^{n+\lambda}) _ { \alpha \beta } \frac{{ \partial w } _ { {ip}}^n}{ \partial x_\beta }  V _ { p } ^ { n }.
Setting in equation \ref{eq:mass_discr} and \ref{eq:stress_discr} into \ref{eq:galerkin} summarizes the space discretization as:
\begin{equation}\label{eq:space_discr}
  \frac{1}{\Delta t} ({m\boldsymbol{v}_i^{n+1}}-{m\boldsymbol{v}_i^{n}}) =
  -\sum _ { p } \boldsymbol{\sigma} _ { p }^{n+\lambda} \nabla w_{ip}^n V _ { p } ^ { n }
  = \boldsymbol{f}_i^{n+\lambda}.
The momentum change of the left side is by construction equal to a (grid node) force.\\
Given that the material point method keeps track of the deformation by a deformation gradient. Each particle will have one associated with it for the deformation of its local neighborhood $\boldsymbol{F}^n_p$. Based on this one may also gain a volume change measure around the particle as $J^n_p = \text{det}(\boldsymbol{F}^n_p)$. Starting with an initial volume of a particle $V_p^0$ the volume may be tracked in time by:
\begin{equation}\label{eq:volume_discr}
  V^n_p \stackrel{\text{\ref{eq:j}}}{\approx} V_p^0J^n_p.
In eq. \ref{eq:piola} an alternate measure for the stress by the first Piola Kirchhoff stress. The results of \ref{eq:space_discr} may therefore equally expressed by it:
  f_{i\alpha} \stackrel{\text{\ref{eq:piola},\ref{eq:volume_discr}}}{=}
  - \sum_p \frac{1}{J^n_p}P^n_{p\alpha\gamma}F^n_{p\beta\gamma}\frac{\partial w_{ip}}{\partial x_\beta}V^0_pJ^n_p
  = - \sum_p P^n_{p\alpha\gamma}F^n_{p\beta\gamma}\frac{\partial w_{ip}}{\partial x_\beta}V^0_p
\cite{MPM:COURSE} \cite{MPM:APIC} \cite{bathe2006finite}
\subsubsection{Deformation gradient evolution}
